Finite Complete Suites for CSP Refinement Testing

### 2018-11-16 scp-journal/article/abstract.tex
revised sections 1 and 2, shortenned paper

In this paper,   new contributions for model-based testing with
Communicating Sequential Processes (CSP) are presented. For a given CSP
process representing the reference model, finite test suites checking the
conformance relations trace refinement and failures refinement are constructed, and
their   completeness (that is, capability to uncover conformity
violations) is proven. The fault domains for which complete failure detection can be guaranteed
are specified by means of normalised transition graphs representing the failures semantics of finite-state CSP processes.
While complete test suites for CSP processes have been
previously investigated by several authors, a sufficient condition for their
finiteness is presented here for the first time. 
Moreover, is shown that the test suites are optimal with respect to two
aspects:~(a)~the maximal length of test traces cannot be further reduced, and (b)~the
nondeterministic behaviour cannot be tested with smaller or fewer sets of
events, without losing the test suite's completeness property.

>>>>

In this paper, new contributions for model-based testing using Communicating
Sequential Processes (CSP) are presented. For a finite non-terminating CSP
process representing the reference model, finite test suites for checking the
conformance relations trace and failures refinement are presented, and their
completeness (that is, capability to uncover conformity violations) is
proven. The fault domains for which complete failure detection can be
guaranteed are specified by means of normalised transition graphs
representing the failures semantics of finite-state CSP processes. While
complete test suites for CSP processes have been previously investigated by
several authors, a sufficient condition for their finiteness is presented
here for the first time. Moreover, it is shown that the test suites are
optimal in two aspects:~(a)~the maximal length of test traces cannot be
further reduced, and (b)~the nondeterministic behaviour cannot be tested with
smaller or fewer sets of events, without losing the test suite's completeness
property.

### 2018-11-16 scp-journal/article/finitecomplete-failures.tex
revised sections 1 and 2, shortenned paper

% -------------------------------------------------------------------------
\subsubsection*{Test Definition and Basic Properties}

In the domain of process algebras, test cases are typically represented by
processes interacting concurrently with the
SUT~\cite{Hennessy:1988:ATP:50497}. Considering an (unknown) process that
represents the behaviour of the SUT, we say that tests synchronise with the
process for the SUT over its visible events and use some additional events
outside the SUT process's alphabet to express whether the test execution
passed or failed.
%%%, or if no verdict could be obtained.

For a given reference process $P$, its normalised transition graph
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
and each integer $j\ge 0$, we define a test for failures refinement as shown
below.
%
\begin{eqnarray}
\label{eq:UFP}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (k < j) \& \big(e:[n]^0   \then U_F(j,k+1,t(n,e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \&  \nonumber
\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(j)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a \emph{test execution} is a trace of the
concurrent process
%\[
$Q\parallel[\Sigma] U_F(j)$.
%\]

It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
occurrence of these events can never be blocked due to a livelock in $Q$
occurring in the same step of the execution.

The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(j)$) if, and
only if, {\it every} execution of $Q\parallel[\Sigma] U_F(j)$ terminates with
the event $\epass$. This can also be  expressed by means of a failures
refinement as defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(j) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test} in the
literature, because every test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}.

We note that it is necessary to use failures refinement in the definition
above, and not just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide
\Sigma$ may have the same visible traces $\varepsilon$ and $\epass$ as the
``Test Passed Process'' $(\epass\then\Stop)$. However, the former may
nondeterministically refuse $\epass$, due to a deadlock occurring when a
faulty SUT process executes concurrently with $U_F(j,k,n)$ executing
branch~(\ref{eq:ufd}), when the guard condition $(k = j\wedge
\minhits(n)\neq\varnothing)$ evaluates to $\ist$. This is explained further
in the next paragraphs. Alternatively, a faulty SUT $Q$ might internally
deadlock after a trace $s$ where $\#s < j$ and $\minhits(G(P)/s) \neq
\varnothing$, so that the process $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks
as well.

Intuitively speaking, $U_F(j)$ is able to perform any trace $s$ of $P$, up to
a length $j$. If, after having already run through $s\in\trc(P)$ with $\#s
\le j$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice. %in the process
%$U_F(j,\#s,n)$ defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
revealed by $\minhits(G(P)/s) = \varnothing$ (recall
(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
external choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also
nondeterministically engage into events in such a situation, branch
(\ref{eq:ufc}) of the test is simultaneously enabled. If $Q/s$ is able to
engage into an event from $\Sigma - [P/s]^0$, a test execution exists where
$U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the $\efail$
event.

If the length of $s$ is still less than $j$, the test accepts any event $e$
from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
$U_F(j,\#s+1,G(P)/s.e)$ (note that $G(P)/s.e = t(G(P)/s,e)$)
in branch~(\ref{eq:ufc}); this follows from
Lemma~\ref{lemma:ufproperties}. A test of this type is called
\emph{adaptive}, because it accepts any legal behaviour of the SUT and adapts
its consecutive behaviour to the event selected by the SUT.

Now suppose that a test execution has run through a trace
$s\in\trc(P)$ of length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n =
G(P)/s$. If $\minhits(n)\neq \varnothing$, the test changes its
behaviour:~instead of offering {\it all} legal events from $[n]^0$ to the
SUT, it nondeterministically chooses a minimal hitting set $H\in \minhits(n)$
and only offers the events contained in $H$. If the SUT refuses to engage
into some event of $H$, this reveals a violation of failures
refinement:~according to Lemma~\ref{lemma:hseta}, a conforming SUT should
accept at least one event of each minimal hitting set in $\minhits(n)$.
Therefore, the test execution only terminates with  $\epass$, if such an
event is accepted by the SUT. Otherwise, it deadlocks, and the test fails.

The specification of $U_F(j,k,n)$ implies that the test always stops after
having engaged into a trace $s\in\trc(Q)$ of maximal length $j$ or $j+1$. If
branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
$j+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
is the last to be entered, the maximal length of $s$ is $j$, and the
execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
entered, the process either accepts another event $e$ of some minimal hitting
set $H\in\minhits(n)$ with $n = G(P)/s$ according to
Lemma~\ref{lemma:ufproperties}. Then the final length of $s$ is $j+1$, and
the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,  the final length of $s$ is $j$,
and the execution stops without a PASS or FAIL event. Such an execution is
also interpreted as FAIL, because it reveals that $(\epass\then\Stop)
\not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma$.

@@

We state basic properties of our tests in the following lemma, to show the
relationships between $U_F(j)$ and the reference process $P$ from which it is
derived.

@@

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that
$$
U_F(j)/s.e = U_F(j,\#s,G(P)/s)/e = (\efail\then\Stop).
$$
%
Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
$P$. \xbox
\end{proof}
%
Note that it is not guaranteed for $U_F(j)$ to run through the traces $s,
s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
some prefix $u$ of $s$: in such a case, $U_F(j)$ may stop with a $\epass$
event by entering branch (\ref{eq:ufb}). Therefore,
Lemma~\ref{lemma:ufproperties} just states the existence of
$U_F(j)$-executions $s, s.e$ satisfying the properties stated there.

% -------------------------------------------------------------------------
\subsubsection*{Complete Testing Assumption}
As explained above, passing a test case $U_F(j)$ requires that none of
the possible executions $(Q\parallel[\Sigma] U_F(j))$ stops after $\efail$ or
stops without having produced event $\epass$. Therefore, it is necessary to determine
whether all possible executions have been covered in the repeated runs of
$(Q\parallel[\Sigma] U_F(j))$.
The theoretical completeness results are therefore based on
a  \emph{complete testing
assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07},
 which means that every
possible behaviour of the SUT is performed after a finite number of test
executions. In practice, this is realised by executing each test several
times, recording the traces that have been performed, and using hardware or
software coverage analysers to determine whether all possible test execution
behaviours of the SUT have been observed. Therefore,
testing nondeterministic SUTs
  comes
at the price of having to apply some grey-box testing techniques enabling us
to decide whether all SUT behaviours have been observed.

@@

A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ counsists of a
reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
over $P$'s alphabet that may or may not conform to $P$.

A test suite $\TS$ is called \emph{complete} with respect to fault model
${\cal F}$, if, and only if, the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}
%
The following main theorem establishes the completeness of our test suite.

% -------------------------------------------------------------------------
\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{$U_F(j)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
% ------------------------------------------------------------------------
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event. This case would also be interpreted as FAIL, since then
$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
U_F(j))\hide \Sigma$.

>>>>

% -------------------------------------------------------------------------
\paragraph{Test Definition and Basic Properties}
In the domain of process algebras, test cases are typically represented by
processes interacting concurrently with the
SUT~\cite{Hennessy:1988:ATP:50497}. Considering an (unknown) process that
represents the behaviour of the SUT, we say that tests synchronise with the
process for the SUT over its visible events and use some additional events
outside the SUT process's alphabet to express whether the test execution
passed or failed.
%%%, or if no verdict could be obtained.

For a given reference process $P$, its normalised transition graph
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
and each integer $j\ge 0$, we define a test for failures refinement as shown
below.
%
\begin{eqnarray}
\label{eq:UFP}
U_F(j) & = & U_F(j,0,\ii n)
\\
U_F(j,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (k < j) \& \big(e:[n]^0   \then U_F(j,k+1,t(n,e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (k = j \wedge \minhits(n)\neq\varnothing) \& \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}

% --------------------------------------------------------------------------
\paragraph{Explanation of the Test Definition} A test \pagebreak is performed by running
$U_F(j)$ concurrently with any SUT process $Q$, synchronising over $\Sigma$.
So, a \emph{test execution} is a trace of the concurrent process
%\[
$Q\parallel[\Sigma] U_F(j)$.
%\]

It is assumed that the events $\efail$ and $\epass$, indicating verdicts FAIL
and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events $\efail$
or $\epass$ always become visible, if they are the only events $U_F(j)/s$ is
ready to engage in: if $U_F(j)/s$ can only produce $\epass$ or $\efail$, the
occurrence of these events can never be blocked due to a livelock in $Q$
occurring in the same step of the execution.

The test is \emph{passed} by the SUT (written $Q\ \pass\ U_F(j)$) if, and
only if, {\it every} execution of $Q\parallel[\Sigma] U_F(j)$ terminates with
the event $\epass$. This can also be  expressed by means of a failures
refinement as defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(j) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}.

It is necessary to use failures refinement in the definition above, and not
just trace refinement:~$(Q\parallel[\Sigma] U_F(j)) \hide \Sigma$ may have
the same visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(j,k,n)$ executing branch~(\ref{eq:ufd}), when
the guard condition $(k = j\wedge \minhits(n)\neq\varnothing)$ evaluates to
$\ist$. This is explained further in the next paragraphs. Alternatively, a
faulty SUT $Q$ might internally deadlock after a trace $s$ where $\#s < j$
and $\minhits(G(P)/s) \neq \varnothing$, so that the process
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks as well.

Intuitively, $U_F(j)$ is able to perform any trace $s$ of $P$, up to a length
$j$. If, after having already run through $s$ with $\#s \le j$, the SUT
accepts an event outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(j)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice. %in the process
%$U_F(j,\#s,n)$ defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice, this is
revealed by $\minhits(G(P)/s) = \varnothing$ (recall
(\ref{eq:minhitminaccempty}) and Lemma~\ref{lemma:ufproperties}). In this
case, the test may terminate successfully (branch (\ref{eq:ufb}) of the
choice in $U_F(j,\#s,G(P)/s)$). If $P/s$ may also nondeterministically engage
into events, branch (\ref{eq:ufc}) is simultaneously enabled. If $Q/s$ is
able to engage into an event in $\Sigma - [P/s]^0$, a test execution exists
where $U_F(j,\#s,G(P)/s)$ branches into (\ref{eq:ufa}) and produces the
$\efail$ event.

If the length of $s$ is still less than $j$, the test accepts any event $e$
from the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as
$U_F(j,\#s+1,G(P)/s.e)$ in branch~(\ref{eq:ufc}); this follows from
Lemma~\ref{lemma:ufproperties}~(note that $G(P)/s.e = t(G(P)/s,e)$). A test
of this type is called \emph{adaptive}, because it accepts any legal
behaviour of the SUT, here any event from $[P/s]^0$, and adapts its
consecutive behaviour to the event selected by the SUT, here
$U_F(j,\#s+1,G(P)/s.e)$.

Now suppose that a test execution has run through a trace $s\in\trc(P)$ of
length $j$, so that $U_F(j)/s = U_F(j,j,n)$ with $n = G(P)/s$. If
$\minhits(n)\neq \varnothing$, the test changes its behaviour:~instead of
offering {\it all} legal events from $[n]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set $H\in \minhits(n)$ and
only offers the events contained in $H$. If the SUT refuses to engage into
some event of $H$, this reveals a violation of failures refinement:~according
to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least one event
of each minimal hitting set in $\minhits(n)$. Therefore, the test execution
terminates with  $\epass$, only if such an event is accepted. Otherwise, it
deadlocks, and the test fails.

The specification of $U_F(j,k,n)$ implies that the test always stops after
having engaged into a trace $s\in\trc(Q)$ of maximal length $j$ or $j+1$. If
branch (\ref{eq:ufa}) is the last to be entered, the maximal length of $s$ is
$j+1$, and the test execution stops with $\efail$. If branch (\ref{eq:ufb})
is the last to be entered, the maximal length of $s$ is $j$, and the
execution stops with $\epass$. If branch (\ref{eq:ufd}) is the last to be
entered, then there are two possibilities. The first is that the process
accepts another event $e$ of some minimal hitting set $H\in\minhits(n)$ with
$n = G(P)/s$ according to Lemma~\ref{lemma:ufproperties}. In this case, the
final length of $s$ is $j+1$, and the execution terminates with $\epass$.
Alternatively, the test execution $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks,
the final length of $s$ is $j$, and the execution stops without a PASS or
FAIL event. Such an execution is also interpreted as FAIL, because it reveals
that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(j)) \hide
\Sigma$.

@@

The following lemma establishes relationships between $U_F(j)$ and the
reference process $P$ from which it is derived.

@@

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that $U_F(j)/s
= U_F(j,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so
$U_F(j,\#s,G(P)/s)$ can engage in $e$ by entering branch (\ref{eq:ufa}). The
specification of this branch implies that $_F(j)/s.e = U_F(j,\#s,G(P)/s)/e =
(\efail\then\Stop)$.

Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because
$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$. Finally, statement
(\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because, according to
(\ref{eq:minhitsGP}), $\minhits(G(P)/s) = \minhits(P/s)$ for all traces of
$P$. \xbox
\end{proof}
%
Note that it is not guaranteed for $U_F(j)$ to run through the traces $s,
s.e$ in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for
some prefix $u$ of $s$: in such a case, $U_F(j)$ may stop with a $\epass$
event by entering branch (\ref{eq:ufb}). Therefore,
Lemma~\ref{lemma:ufproperties} just states the existence of
$U_F(j)$-executions $s, s.e$ satisfying the properties stated there.

% -------------------------------------------------------------------------
\paragraph{Complete Testing Assumption} As explained above, passing a test case
$U_F(j)$ requires that none of the possible executions $(Q\parallel[\Sigma]
U_F(j))$ stops after $\efail$ or stops without having produced the event
$\epass$. Therefore, it is necessary to determine whether all possible
executions have been covered in the repeated runs of $(Q\parallel[\Sigma]
U_F(j))$. The theoretical completeness results are, therefore, based on a
\emph{complete testing
assumption}~\cite{hierons_testing_2004,DBLP:conf/icfem/CavalcantiG07}, which
means that every possible behaviour of the SUT is performed after a finite
number of test executions. In practice, this is realised by executing each
test several times, recording the traces that have been performed, and using
hardware or software coverage analysers to determine whether all possible
behaviours of the SUT have been observed. Therefore, testing nondeterministic
SUTs comes at the price of having to apply some grey-box testing techniques.
%enabling us to decide whether all SUT behaviours have been observed.

@@

A CSP \emph{fault model} ${\cal F} = (P,\sqsubseteq,{\cal D})$ consists of a
reference process $P$, a conformance relation $\sqsubseteq \in \{\lessdet_T,
\lessdet_F\}$, and a fault domain ${\cal D}$, which is a set of CSP processes
over $P$'s alphabet that may or may not conform to $P$. A test suite $\TS$ is
called \emph{complete} with respect to fault model ${\cal F}$, if, and only
if, the following conditions are fulfilled.
\begin{description}
\item[1.~Soundness] If $P \sqsubseteq Q$, then $Q$ passes all tests in $\TS$.
\item[2.~Exhaustiveness] If $P \not\sqsubseteq Q$ and $Q\in{\cal D}$,
then $Q$ fails at least one test in $\TS$.
\end{description}
%
The following main theorem establishes the completeness of our test suite.

% -------------------------------------------------------------------------
\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(j)~|~0 \le j < pq  \}\quad\text{with $U_F(j)$ as specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
% ------------------------------------------------------------------------
%
The proof of the theorem follows directly from the two lemmas below. The
first lemma establishes that test suite $\TS_F$ is sound, and the second
establishes that the suite is also exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
A test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}We make two points in separate steps below. The
first is that the test execution cannot reach branch~(\ref{eq:ufa}) and raise
a $fail$ event.  The second is that it cannot deadlock without raising a
$pass$ event. This case would also be interpreted as FAIL, since then
$\epass\then\Stop$ is not failures refined by $(Q\parallel[\Sigma]
U_F(j))\hide \Sigma$.
